\documentclass[openany, 12pt]{article}
\input{preamble}
\title{Enhanced Flank Steak with Butter, Onions, and Garlic}
\author{Idris}
\date{May 2025}

%chktex-file 1
% chktex-file 2999999
\begin{document}

\maketitle
\tableofcontents

\section{Ingredients}
\begin{itemize}
	\item 1 flank steak (already cooked via sous vide)
	\item A generous amount of butter (at least 4 tablespoons, adjust to your preference)
	\item Several onions (2-3 medium), sliced
	\item Several cloves of garlic (3-5), minced
	\item Salt and freshly ground black pepper to taste
	\item Optional: Red wine, beef broth, or balsamic vinegar (for pan sauce variation)
	\item Optional: Fresh herbs (thyme, rosemary, parsley), chopped (for infused butter or garnish)
	\item Optional: Cooking oil (if needed for searing)
\end{itemize}

\section{Tools}
\begin{itemize}
	\item Stainless steel skillet (heavy-bottomed preferred)
	\item Cutting board
	\item Sharp knife
	\item Paper towels
	\item Spoon
	\item Small saucepan (for infused butter, optional)
\end{itemize}

\section{Steps}
\subsection{Simple and Classic Sauté (Stainless Steel Pan)}
\begin{enumerate}
	\item Pat the sous vide flank steak completely dry with paper towels.
	\item Place the stainless steel skillet over medium-high heat. Once the pan
	      is hot, add a generous amount of butter.
	\item Add the minced garlic to the pan and cook until fragrant (about 1
	      minute). Be careful not to burn the garlic. You can remove some of the
	      onion-garlic mixture and set it aside for topping later if desired.
	\item Carefully place the dried flank steak in the hot skillet. Sear for 1-2
	      minutes per side until a deep brown crust forms. Avoid overcrowding the
	      pan; if necessary, sear in batches.
	\item \textbf{Optional Basting:} During the last 30 seconds of searing,
	      return the reserved onion-garlic mixture (if any) and any accumulated
	      juices from the sous vide bag to the pan. Tilt the pan and use a spoon
	      to continuously baste the steak with the flavorful butter and aromatics.
	\item Remove the steak from the pan and transfer it to a clean cutting board.
	\item Let the steak rest for 5-10 minutes before slicing thinly against the grain.
	\item Spoon the pan sauce (butter, onions, garlic, and juices) over the
	      sliced steak and serve immediately.
\end{enumerate}

\subsection{Flavorful Pan Sauce Variation (Stainless Steel Pan)}
\begin{enumerate}
	\item Follow steps 1-6 from the Simple and Classic Sauté method.
	\item Remove the steak from the pan and set it aside to rest. Reduce the
	      heat to medium.
	\item \textbf{Deglaze the pan:} Add a splash of red wine, beef broth, or
	      balsamic vinegar to the hot skillet. Use a wooden spoon or spatula to
	      scrape up any browned bits from the bottom of the pan. \item Stir in the
	      cooked onions and garlic (or add fresh minced garlic and cook for 30
	      seconds). Let the liquid reduce slightly, concentrating the flavors.
	\item Swirl in a knob of fresh butter at the end for richness and shine.
	      Season the sauce with salt and pepper to taste. \item Slice the rested
	      steak thinly against the grain and spoon the flavorful pan sauce
	      generously over it. \end{enumerate}

\subsection{Garlic Herb Butter Infusion (Optional)} \begin{enumerate}
	\item While the steak is resting, melt a good amount of butter in a
	      small saucepan over low heat. \item Finely mince fresh garlic
	      and stir it into the melted butter. Add chopped fresh herbs if
	      using. \item Let the garlic and herbs infuse the butter for a
	      few minutes. \item Slice the rested steak and drizzle the fragrant
	      garlic herb butter over the top. \end{enumerate}

\section{Timing} \begin{itemize} \item Prep Time: Approximately 5-10 minutes
	      (mostly slicing onions and mincing garlic) \item Sautéing Onions: 5-7
	      minutes \item Sautéing Garlic: Approximately 1 minute \item Searing Steak:
	      2-4 minutes total (1-2 minutes per side) \item Resting Steak: 5-10 minutes
	\item Pan Sauce (if making): 5-7 minutes \item Garlic Herb Butter (if
	      making): 5 minutes (during steak rest) \item Total Time (approximate):
	      15-30 minutes (depending on the chosen method and resting time)
\end{itemize}

\end{document}

% \printbibliography{} \printindex{}
\end{document}
